%----------------------------------------------------------------------------------------
%	ACKNOWLEDGEMENTS
%----------------------------------------------------------------------------------------

\begin{acknowledgements}
    \addchaptertocentry{\acknowledgementname} % Add the acknowledgements to the table of contents
    
    Crikey, I never thought I'd get to this point, but here we are, it’s finally done! Time really does fly when you're having fun, or when you have a hundred little fiddles you want to try on your data that might just squeeze out just that little bit more interesting science... No PhD is a solitary endeavour, and so there are a number of important people to whom I'd like to say thank you for helping this come to fruition.

    First and foremost, the largest amount of thanks must of course go to Patrick. Patrick has been key in getting me into pulsar research in the first place -- from the summer project in 2016, through the MPhys project, and now on to this PhD, Patrick has been incredibly supportive and patient over the entire process. I’ll never forget sitting eating dragon fruit with you in a thunderstorm on the Great Wall of China, or hacking through the jungle around FAST! Thank you for sparking my interest in this field, and for your fantastic insights, support, and sage wisdom throughout.
        
    It has been an absolute pleasure to be amongst the amazing people of the wonderful JBCA postgraduate community, both in the office and in more recent times on the Discord server. Our ridiculous conversations over PGTea, in the virtual office, and general chatter have been the high points of my academic life, and the support available has been second to none. Special credit to Laura for putting together this amazing thesis template! I'm sure we will all remain in contact for many years to come.
      
    I would also like to say thank you to the folks at the NAOC and FAST who welcomed me and Patrick during our visit in 2018. What an experience that was! In particular, Di Li, Lei Zhang, and Yi Feng, whose work and friendship were key to a lot of the results in this thesis. Special thanks also go to Geoff Wright for many enthusiastic conversations on pulsar geometry, aliasing, and carousels over the last four years.
    
    The Squad (you know who you are) must also be acknowledged for keeping me sane (albeit to a questionable standard) through the many fun times we've shared. From our weekends away, to the Commune, to Banff and Taskmaster Zoom calls during the pandemic, you are the best friendship group I could have asked for.
    
    Last but by no means least, Tilly. You have been an absolute rock over the last few years, and I genuinely don't think I could have done this without all your support and encouragement. Here's to many more amazing adventures, and nonsense late night conversations! I can only hope to be there as much for you when your thesis time comes. Du \"ar den b\"asta av alla \"agg.
    
    \end{acknowledgements}
    
    