
\begin{declaration}
    \addchaptertocentry{\authorshipname} % Add the declaration to the table of contents
    \noindent {\bcf YOU WILL NEED TO CHANGE THIS. Check the official requirements for declarations: \href{https://www.bmh.manchester.ac.uk/doctoral-academy/your-phd/thesis-submission/presentation-regulations/}{University of Manchester Doctoral Presentation Regulations}}
    
    \noindent I, \authorname, declare that this thesis titled, \enquote{\ttitle} and the work presented in it are my own. I confirm that:
    
    %% You WILL NEED TO CHANGE THIS!!! Go here to find out how:
    %% https://www.bmh.manchester.ac.uk/doctoral-academy/your-phd/thesis-submission/presentation-regulations/
    
    \begin{itemize} 
    \item This work was done wholly or mainly while in candidature for a research degree at this University.
    \item Where any part of this thesis has previously been submitted for a degree or any other qualification at this University or any other institution, this has been clearly stated.
    \item Where I have consulted the published work of others, this is always clearly attributed.
    \item Where I have quoted from the work of others, the source is always given. With the exception of such quotations, this thesis is entirely my own work.
    \item I have acknowledged all main sources of help.
    \item Where the thesis is based on work done by myself jointly with others, I have made clear exactly what was done by others and what I have contributed myself.\\
    \end{itemize}
     
    % \noindent Signed:\\
    % \rule[0.5em]{25em}{0.5pt} % This prints a line for the signature
     
    % \noindent Date:\\
    % \rule[0.5em]{25em}{0.5pt} % This prints a line to write the date
\end{declaration}
    
    % \cleardoublepage
    
    %----------------------------------------------------------------------------------------
    %	COPYRIGHT STATEMENT PAGE
    %----------------------------------------------------------------------------------------
    %% These exact words are required by UoM in this exact spot.
    
    \begin{copyrightstatement}
    
    \addchaptertocentry{\copyrightname} % Add the declaration to the table of contents
    \noindent {\bcf This page is a requirement, and must contain the exact words required by UoM. To double check that these words are up-to-date, go here: \href{https://www.bmh.manchester.ac.uk/doctoral-academy/your-phd/thesis-submission/presentation-regulations/}{University of Manchester Doctoral Presentation Regulations}}
    
    \begin{enumerate}[label=(\roman*)]
    \item The author of this thesis (including any appendices and/or
    schedules to this thesis) owns certain copyright or related rights
    in it (the ``Copyright'') and s/he has given The University of
    Manchester certain rights to use such Copyright, including for
    administrative purposes.\\[0.3cm]
    \item Copies of this thesis, either in full or in extracts and whether in
    hard or electronic copy, may be made {\bcf only} in accordance with the
    Copyright, Designs and Patents Act 1988 (as amended) and
    regulations issued under it or, where appropriate, in accordance
    with licensing agreements which the University has from time to
    time. This page must form part of any such copies made.\\[0.3cm]
    \item The ownership of certain Copyright, patents, designs, trademarks
    and other intellectual property (the ``Intellectual Property'') and
    any reproductions of copyright works in the thesis, for example
    graphs and tables (``Reproductions''), which may be described in
    this thesis, may not be owned by the author and may be owned
    by third parties. Such Intellectual Property and Reproductions
    cannot and must not be made available for use without the prior
    written permission of the owner(s) of the relevant Intellectual
    Property and/or Reproductions.\\[0.3cm]
    \item Further information on the conditions under which disclosure,
    publication and commercialisation of this thesis, the Copyright
    and any Intellectual Property and/or Reproductions described in
    it may take place is available in the University IP Policy (see
    \href{http://documents.manchester.ac.uk/DocuInfo.aspx?DocID=2442
    0}{documents.manchester.ac.uk}), in any relevant Thesis restriction declarations deposited in the
    University Library, The University Library’s regulations (see
    \href{http://www.library.manchester.ac.uk/about/regulations/}{www.library.manchester.ac.uk/about/regulations/}) and in
    The University’s policy on Presentation of Theses
    \end{enumerate}
    \end{copyrightstatement}
    
    % \cleardoublepage
    